---
title: "Jul.6"
author: "Xinyu Zhang"
date: "7/6/2020"
output:
  pdf_document: default
---
## Generating the Files
```{r}
library(dplyr)
#reading in listening test results:
listener1 <- read.csv("Results_AK.csv", header = TRUE, sep = ";")
listener2 <- read.csv("Results_KS.csv", header = TRUE, sep = ";")

#renaming the two answers, removing the column "Num", 
#and adding a dummy variable (= speaker + T) as an identifier for when joining two datasets:
SLP1 <- listener1 %>%
  dplyr::select(-Num) %>% 
  rename(Roughness = Answer1, Breathiness = Answer2)
SLP1$dummy <- paste(SLP1$Speaker, "-", SLP1$T)
#nrow(SLP1) #96
SLP2 <- listener2 %>%
  rename(Roughness = Answer1, Breathiness = Answer2)%>%
  dplyr::select(-Num)
SLP2$dummy <- paste(SLP2$Speaker, "-", SLP2$T)
#nrow(SLP2) #96

AVQI <- read.csv("AVQI_results_new.csv")
#adding an dummy variable (= speaker + T) as an identifier when joining two datasets:
AVQI$dummy <- paste(AVQI$Speaker, "-", AVQI$T)
#converting "Jitter" as a numeric value from a factor (this introduces new NAs):
AVQI$Jitter <- as.numeric(levels(AVQI$Jitter))[AVQI$Jitter]

#defining the mean and sd of the Brethiness and Roughness scores:
library(dplyr)
mB1 <- mean(SLP1$Breathiness)#AK
sdB1 <- sd(SLP1$Breathiness)#AK
mB2 <- mean(SLP2$Breathiness)#KS
sdB2 <- sd(SLP2$Breathiness)#KS
mR1 <- mean(SLP1$Roughness)#AK
sdR1 <- sd(SLP1$Roughness)#AK
mR2 <- mean(SLP2$Roughness)#KS
sdR2 <- sd(SLP2$Roughness)#KS

#adding a column containing z-values for each NLP score table
zSLP1 <- SLP1 %>% 
  mutate(zBreathy = (Breathiness-mB1)/sdB1, zRough = (Roughness-mR1)/sdR1) %>% 
  select(-A, -Z, -T)
zSLP2 <- SLP2 %>% 
  mutate(zBreathy = (Breathiness-mB2)/sdB2, zRough = (Roughness-mR2)/sdR2) %>% 
  select(-A, -Z, -T)

#joining the scores from two SLPs, by matching the dummy identifier, and removing redundant rows
zSLP <- full_join(zSLP1, zSLP2, by = "dummy") %>% 
  select(-Speaker.x, -Speaker.y)

#calculating the mean of z scores for each token of recording
zSLP$mzB = rowMeans(zSLP[,c('zBreathy.x', 'zBreathy.y')], na.rm=TRUE)
zSLP$mzR = rowMeans(zSLP[,c('zRough.x', 'zRough.y')], na.rm=TRUE)

#join the two SLP results tables with AVQI by matching both Speaker and T
zAVQI_ <-inner_join(AVQI,zSLP, by = "dummy")
zAVQI <- na.omit(zAVQI_)#removing rows containing NAs

#write the file for later convenience
write.csv(zAVQI, "zAVQI.csv")
#zAVQI.csv now has the AVQI data and the SLP ratings combined and matched, 
#and with the variables in the right classes
library(ggplot2)
ggplot(data = zAVQI, aes(x = mzB, y = mzR)) + 
  geom_point() +
  geom_smooth(method = "lm")
#ggsave("z_Corre.png", plot = last_plot(), device = NULL, scale = 1, width = 10, height = 6, units = c("cm"), dpi = 600)
```

## Step-wise model comparison
```{r}
#decorrelating breathiness from roughness:
breathyRough <- lm(mzB ~ mzR, data = zAVQI)
plot(breathyRough)
mzBCorr <- fitted(breathyRough, zAVQI$mzR)
zAVQI$mzBDecorr <- (zAVQI$mzB - mzBCorr)

#using step to determine the best linear model
##roughness
zRoughModel.tmp <- lm(mzR ~ CPPS + HNR + Jitter + Shimmer + ShdB + Slope + Tilt, zAVQI)
#using BIC instead of AIC
zRoughModel <- step(zRoughModel.tmp, k = log(nrow(zAVQI)))

##breathiness
zBrDecorrModel.tmp <- lm(mzBDecorr ~ CPPS + HNR + Jitter + Shimmer + ShdB + Slope + Tilt, zAVQI)
#zBrDecorrModel <- step(zBrDecorrModel.tmp)
#using BIC instead of AIC:
zBrDecorrModel <- step(zBrDecorrModel.tmp, k = log(nrow(zAVQI)))

#summarizing the two models
summary(zRoughModel)
summary(zBrDecorrModel)

#mBreathyModel <- step(mBreathyModel.tmp)
#mBreathyModel <- step(mBreathyModel.tmp, k=log(nrow(zAVQI)))
#AIC: k=2
#BIC: k=log(n)
#BIC includes as few predictors as possible, avoids overfitting
```

^The Above suggests that the best predictors for the roughness model are CPPS and HNR, ShdB and Tilt; the best predictors for the decorrelated breathiness model are CPPS, HNR, and ShdB, Shimmer.

Sep15: best predictors for roughness are HNR and the spectral tilt. 
best predictors for decoorrelated breathiness are CPPS and HNR.

Since the variables for two models overlap, the rebuilt (correleated-to-roughness) breathiness model will use exactly the same variables as the decorrelated model, only with the original (non-decorrelated) breathiness scores as the independent variable.

## Generating the output
```{r}
#running the models:
rough.model <- lm(mzR ~ HNR + Tilt, data = zAVQI)
summary(rough.model)

breathyDecorr.model <- lm(mzBDecorr ~ HNR + CPPS, data = zAVQI)
summary(breathyDecorr.model)

#the reconstructed breathiness model 
#(using "mzB" instead of "mzBDecorr" as the independent variable here:
breathy.model <- lm(mzB ~ HNR + CPPS + Tilt, data = zAVQI)
summary(breathy.model)

# mBreathyModel #<- step(mBreathyModel.tmp)
# mBreathyModel <- step(mBreathyModel.tmp, k=log(nrow(zAVQI)))
#AIC: k=2
#BIC: k=log(n)

#https://www.biorxiv.org/content/biorxiv/early/2019/01/08/449751.full.pdf

# radiochemo (improves better) and laser therapy severe 

#tumors on the vocal folds 
#laser ecxitions or radio therapy or both because ethical committee approval not for medical
#p-value not as useful anymore because the data has already undergone model selection
#look at adjusted R-squared instead
plot(breathyDecorr.model$residuals)
```
Both CPPS and HNR are significant for Roughness (Adjusted R-squared = 0.3243). Only CPPS turned out to be significant for Breathiness (not including ShdB could result in a better p-value for HNR, this could be argued by adding in one more step to evaluate the BreathDecorr model), Adjusted R-squared = 0.5692.

## Plots
```{r}
library(scatterplot3d)
par(mfrow = c(1,2))
#roughness:
rough3d<- scatterplot3d(zAVQI$HNR, zAVQI$CPPS, zAVQI$mzR, xlab = "HNR", ylab = "CPPS", zlab = "Roughness-Z")
rough3d$plane3d(rough.model)

#breathiness:
breathyDecorr3d<- scatterplot3d(zAVQI$HNR, zAVQI$CPPS, zAVQI$mzB, xlab = "HNR", ylab = "CPPS", zlab = "Breathiness-Z(Decorrelated)")
breathyDecorr3d$plane3d(breathyDecorr.model)

```

```{r}
#linear regression
head(zAVQI)
library(lme4)
#setting contrast
contrast <- cbind(c(0.5, -0.5)) #AK,KS
colnames(contrast) <- c("SubjectAK-SubjectKS")
contrasts(table$Subject) <- contrast
#Roughness ~ Breathiness, with Speaker as a random effect
model1 <- lmer(Roughness ~ Subject + (Subject | Speaker), data = table, REML = TRUE, na.action = "na.omit")

model2 <- lm(Roughness ~ Subject, data = zAVQI, REML = TRUE, na.action = "na.omit")

model3 <- lm(Breathiness ~ Subject, data = table, REML = TRUE, na.action = "na.omit")

summary(model2)
```
```{r}
ggplot(data = zAVQI, aes(x = mzR, y = mzB)) + 
  geom_point() +
  geom_smooth(method = "lm")
ggsave("z_Corre.png", plot = last_plot(), device = NULL)
```

```{r}
ggplot(data = zAVQI, aes(x = mzR, y = mzBDecorr)) + 
  geom_point() +
  geom_smooth(method = "lm")
#ggsave("rough-DecorrBreathy.png", plot = last_plot(), device = NULL)
```
