---
title: "linear models"
author: "Xinyu Zhang"
date: "6/26/2020"
output: html_document
---
Reshaping the tables:
```{r}
library(dplyr)
#reading in listening test results
listener1 <- read.csv("Results_AK.csv", header = TRUE, sep = ";")
listener2 <- read.csv("Results_KS.csv", header = TRUE, sep = ";")
#renaming the two answers, removing column "Num"
SLP1 <- listener1 %>%
  dplyr::select(-Num) %>% 
  rename(Roughness = Answer1, Breathiness = Answer2)
SLP1$dummy <- paste(SLP1$Speaker, "-", SLP1$T)
#nrow(SLP1) #96

SLP2 <- listener2 %>%
  rename(Roughness = Answer1, Breathiness = Answer2)%>%
  dplyr::select(-Num)
SLP2$dummy <- paste(SLP2$Speaker, "-", SLP2$T)
#nrow(SLP2) #96

AVQI <- read.csv("AVQI_results_new.csv")
AVQI$dummy <- paste(AVQI$Speaker, "-", AVQI$T)
AVQI$Jitter <- as.numeric(levels(AVQI$Jitter))[AVQI$Jitter]

AVQI_RB_1 <- inner_join(SLP1,AVQI, by = "dummy")#retain only rows in both sets
#nrow(AVQI_BR_1) #96

AVQI_RB_2 <- inner_join(SLP2, AVQI, by = "dummy")#retain only rows in both sets
#nrow(AVQI_BR_2) #96

#join the two listening results tables by matching both Speaker and T
AVQI_RB <-rbind(AVQI_RB_1,AVQI_RB_2) %>% 
  dplyr::select(-A, -Speaker.x, -Speaker.y, -T.x, -T.x, -T.y, -Version, -Z)
#nrow(AVQI_RB) #192 (= 96 *2)
glimpse(AVQI_RB)
```
The Breathiness and Roughness ratings are treated as separate instead of averaged over two SLPs

The Averaged alternative:
```{r}
mSLP <-full_join(SLP1,SLP2, by = "dummy") %>% 
  dplyr::select(-T.x, -Speaker.y, -Z.x, -Z.y, -T.y, -Speaker.x, -A.x, -A.y)

mSLP$RoughMean <- rowMeans(mSLP[,c('Roughness.x', 'Roughness.y')], na.rm=TRUE)

mSLP$BreathyMean <- rowMeans(mSLP[,c('Breathiness.x', 'Breathiness.y')], na.rm=TRUE)

mSLP <- mSLP %>% 
  dplyr::select(dummy, RoughMean, BreathyMean)

#nrow(mSLP) #96 

AVQI_mRB <- inner_join(mSLP,AVQI, by = "dummy") %>% 
  dplyr::select(-Speaker, -T, -Version) %>% 
  glimpse
```

On averaged data:
```{r}
mBreathyModel <- lm(BreathyMean ~ RoughMean, AVQI_mRB)
summary(mBreathyModel)
```

```{r}
mBreathyCorrelated <- fitted(mBreathyModel, AVQI_mRB$RoughMean)

# Part of Breathiness not predicted by Roughness
AVQI_mRB$mBreathyDecorr <- AVQI_mRB$BreathyMean - mBreathyCorrelated
str(AVQI_mRB)
```

```{r}
AVQI_mRB_c <- na.omit(AVQI_mRB)#removing NAs
mRoughModel.tmp <- lm(RoughMean ~ CPPS + HNR + Jitter + Shimmer + ShdB + Slope + Tilt, AVQI_mRB_c)
step(mRoughModel.tmp)
mRoughModel <- step(mRoughModel.tmp)
summary(mRoughModel)
```
CPPS and HNR

```{r}
library(MASS)
AVQI_mR<- AVQI_mRB_c %>% 
  dplyr::select(-dummy, -BreathyMean, -mBreathyDecorr, -AVQI)
min.modelRm <- lm(RoughMean ~., data = AVQI_mR)
step.modelRm<- stepAIC(min.modelRm, direction = "both", 
                      trace = FALSE)
summary(step.modelRm)
```
CPPS and HNR

```{r}
mRoughModel <- lm(RoughMean ~ CPPS + HNR, AVQI_mR)
summary(mRoughModel)
```

For Breathiness:
```{r}
mBreathyModel.tmp <- lm(mBreathyDecorr ~ CPPS + HNR + Jitter + Shimmer + ShdB + Slope + Tilt, AVQI_mRB_c)
step(mBreathyModel.tmp)
mBreathyModel <- step(mBreathyModel.tmp)
summary(mBreathyModel)
```
CPPS and HNR

Alternative:
```{r}
AVQI_mB<- AVQI_mRB_c %>% 
  dplyr::select(-dummy, -RoughMean, -BreathyMean, -AVQI)
min.modelB <- lm(mBreathyDecorr ~., data = AVQI_mB)
step.modelB<- stepAIC(min.modelB, direction = "both", 
                      trace = FALSE)
summary(step.modelB)
```
Also CPPS and HNR

```{r}
mBreathyModel <- lm(mBreathyDecorr ~ CPPS + HNR, AVQI_mB)
summary(mBreathyModel)
```

Trying on un-averaged data:
```{r}
BreathyModel <- lm(Breathiness ~ Roughness, AVQI_RB)
summary(BreathyModel)
```

```{r}
BreathyCorrelated <- fitted(BreathyModel, AVQI_RB$Roughness)

# Part of Breathiness not predicted by Roughness
AVQI_RB$BreathyDecorr <- AVQI_RB$Breathiness - BreathyCorrelated


```

For Roughness:
```{r}
#removing NAs
AVQI_RB_c <- na.omit(AVQI_RB)
RoughModel.tmp <- lm(Roughness ~ CPPS + HNR + Jitter + Shimmer + ShdB + Slope + Tilt, AVQI_RB_c)
step(RoughModel.tmp)
RoughModel <- step(RoughModel.tmp)
summary(RoughModel)
```
CPPS, HNR, Tilt


```{r}
AVQI_R<- AVQI_RB_c %>% 
  dplyr::select(-dummy, -Breathiness, -BreathyDecorr, -AVQI, -Subject)
min.modelR <- lm(Roughness ~., data = AVQI_R)
step.modelR<- stepAIC(min.modelR, direction = "both", 
                      trace = FALSE)
summary(step.modelR)
```
Also CPPS, HNR,Tilt

```{r}
RoughModel <- lm (Roughness ~ CPPS + HNR + Tilt, data = AVQI_R)
summary(RoughModel)
```


For Breathiness:
```{r}
BreathyModel.tmp <- lm(BreathyDecorr ~ CPPS + HNR + Jitter + Shimmer + ShdB + Slope + Tilt, AVQI_RB_c)
step(BreathyModel.tmp)
BreathyModel <- step(BreathyModel.tmp)
summary(BreathyModel)
```
CPPS, HNR, ShdB

Alternative:
```{r}
AVQI_B<- AVQI_RB_c %>% 
  dplyr::select(-dummy, -Roughness, -Breathiness, -AVQI, -Subject)
min.modelB <- lm(BreathyDecorr ~., data = AVQI_B)
step.modelB<- stepAIC(min.modelB, direction = "both", 
                      trace = FALSE)
summary(step.modelB)
```
Also CPPS, HNR, ShdB

```{r}
BreathyModel <- lm(BreathyDecorr ~ CPPS + HNR + ShdB, AVQI_B)
summary(BreathyModel)
```
ShdB not significant.

Combining the correlated part:
```{r}
BreathyModel <- lm(Breathiness ~ CPPS + HNR + Tilt + ShdB, data = AVQI_RB)
summary(BreathyModel)
```
ShdB not significant. Adjusted R-squared: 0.5176
